---
title: "Reproductible research - Project 1"
author: "Beatriz Del Fiol"
date: "16 de junho de 2019"
output: html_document
---

1 - Code for reading in the dataset and/or processing the data

```{r 1chuck}
setwd("C:/Users/beatr/Desktop/CourseraR/5_Reproducible research/Project 1")
activity <- read.csv("C:/Users/beatr/Desktop/CourseraR/5_Reproducible research/Project 1/activity.csv")
activity <- read.csv("activity.csv")
head(activity)
str(activity)
steps <- as.numeric("steps")
date <- as.character.Date(date)
class(date)
sum(is.na(activity))
#na.omit(activity)


library(magrittr)
library(dplyr)
library(lubridate)
```

2 - Histogram of the total number of steps taken each day

```{r 2chunck}

stepsperday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
#class(date)
#class(steps)
head(stepsperday)
date  <- as.Date(date, "%Y-%m-%d")
hist(stepsperday$steps, xlab = "Steps per day", ylab = "Frequency", main = "Histogram: Steps per Day", breaks = 10)

```

3 - Mean and median number of steps taken each day

```{r 3chunck}
meansteps <- mean(stepsperday$steps, na.rm = TRUE)
meansteps

mediansteps <- median(stepsperday$steps, na.rm = TRUE)
mediansteps

summary(stepsperday)
```

4 - The 5-minute interval that, on average, contains the maximum number of steps

```{r 4chunck}

library(magrittr)
library(dplyr)

stepsperinterval <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
#head(stepsperinterval)
plot(x = stepsperinterval$interval, 
     y = stepsperinterval$steps, 
     type = "l", 
     col = "blue",
     xlab = "Intervals",
     ylab = "Average Steps Taken per Days",
     main = "Daily Activity Pattern")

intervalmaxsteps <- stepsperinterval$interval[which.max(stepsperinterval$steps)]
intervalmaxsteps

```

5 - Code to describe and show a strategy for imputing missing data - Using mean for that day to replace the missing data

```{r 5chunck}
missingVals <- sum(is.na(activity))
#missingVals

activityNA <- activity
head(activityNA)
nas <- is.na(activityNA$steps)

avg_interval <- tapply(activityNA$steps, activityNA$interval, mean, na.rm=TRUE, simplify = TRUE)
head(avg_interval)
activityNA$steps[nas] <- avg_interval[as.character(activityNA$interval[nas])]
names(activityNA)
head(activityNA)

stepsperday <- aggregate(steps ~ date, data = activity, FUN = sum)
str(stepsperday)
date  <- as.Date(date, "%Y-%m-%d")
head(stepsperday)

stepsperintervalNA <- aggregate(steps ~ interval, data = activityNA, FUN = mean, na.rm = TRUE)
head(stepsperintervalNA)

```

6 - Histogram of the total number of steps taken each day after missing values are imputed

```{r 6chunck}

stepsperdayNA <- aggregate(steps ~ date, data = activityNA, FUN = sum, na.rm = TRUE)
head(stepsperdayNA)
date  <- as.Date(date, "%Y-%m-%d")

hist(stepsperdayNA$steps, xlab = "Steps per day", ylab = "Frequency", main = "Histogram: Steps per Day", breaks = 10)

```

7 - Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r 7chunck}

activityNA$date <- as.Date(activityNA$date)
#wday(activityNA$date, label=TRUE)
activityNA$weekday = weekdays(activityNA$date)
head(activityNA)
activityNA$DayCategory <- ifelse(activityNA$weekday %in% c("sábado", "domingo"), "Weekend", "Weekday")
head(activityNA)


library(ggplot2)
ggplot(activityNA, aes(x =interval , y=steps, color=DayCategory)) +
        geom_line() +
        labs(title = "Ave Daily Steps Day Category", x = "Interval", y = "Total Number of Steps") +
        facet_wrap(~ DayCategory, ncol = 1, nrow=2)


```







